---
title: "RNA viruses -Narnaviridae"
date: "February 26, 2019"
author: "Erin Nuccio with input from Evan Starr"
output: html_document
---


Load libraries and set working directory
```{r}
library("DESeq2")
library("Biobase")
library("ggplot2")
library("data.table")
library("pheatmap")
library("dplyr")
library("tibble")
library("vegan")

working_directory = "/Users/starr/Documents/papers/rna_virus/R work/Euk Viruses/"
setwd(working_directory)

```

  #########
  ### Import data
  #########

```{r}

infile = "Reads_coverage_Euk_Viruses_fromwheeltree-Narna.txt"
mtx_counts_N = read.table(infile, header = TRUE, row.names = 1,  sep="\t")

mtx_matrix_N = as.matrix(sapply(mtx_counts_N, as.integer)) 
rownames(mtx_matrix_N) = rownames(mtx_counts_N)
mtx_matrix_N[is.na(mtx_matrix_N)] = 0
mtx_matrix_N = mtx_matrix_N[,order(colnames(mtx_matrix_N))]

designfile = "mtx_design_ecofun_genomes4.txt" 
mtx_design = read.delim(designfile, header = TRUE, row.names = 1)
mtx_design <- mtx_design[!(rownames(mtx_design) %in% c("H3_Rhizo_39")),] # Remove H3_Rhizo_39 because these columns are mostly 0s, was messing up the geometric mean 
# Keeping "H2_BulkLitter_50" for now
design_rownames = rownames(mtx_design)
```

  #########
  ### Create dataset
  #########  

```{r}
dds = DESeqDataSetFromMatrix(countData = mtx_matrix_N, colData = mtx_design, design = ~ microcosm)
dds = dds[ rowSums(counts(dds)) > 5, ] # Filter out rows with < 5 counts
  
### Create custom groups
dds$group = factor(paste0(dds$treatment, dds$location))
dds$group_harvest = factor(paste0(dds$group, dds$harvest))
dds$group_harvest_location = factor(paste0(dds$harvest, dds$location))
dds$group_harvest_treatment = factor(paste0(dds$harvest, dds$treatment))
dds$group_htl = factor(paste0(dds$harvest, dds$treatment, dds$location))
  #dds$microcosm <- droplevels(dds$microcosm)
  
### Analyze results
design(dds) = ~ group_htl
dds = DESeq(dds)
res = results(dds)
resultsNames(dds)

dds_counts = counts(dds, normalized=TRUE)
write.table(dds_counts, file = paste(working_directory,"/DESeq_normalized_euk_virus-Narna.txt", sep = ""))
saveRDS(dds, file = paste(working_directory,"/dds_htl_euk_virus-Narna.rds", sep = ""))
```


Make a quick PCA in QIIME

```{bash}
source /macqiime/configs/bash_profile.txt

# Formated text file in Excel before converting to biom (made tab delimited, added #OTU_ID to first header)
biom convert -i  DESeq_normalized_euk_virus-narna.biom  -o DESeq_normalized_euk_virus_qiime-narna.txt --to-json --table-type="OTU table"

beta_diversity_through_plots.py -i DESeq_normalized_euk_virus_qiime-narna.biom -o beta_diversity_euk_virus-Narna -m mapping_mtx_design_ecofun_genomes4.txt -p parameter.txt -f

```

###### 1. metaMDS to create ordination object

Run Adonis (PERMANOVA)

```{r}
library("vegan")

# The dds_dist was created by QIIME
  dds_bray_dm_N = read.delim("beta_diversity_euk_virus-narna/bray_curtis_dm.txt",header = TRUE, row.names = 1)
dds_dist_N = as.dist(dds_bray_dm_N)

dds_metaMDS_N<-metaMDS(dds_dist_N, distance = "bray", k = 2, trymax = 1000, autotransform =FALSE, noshare = 0.1, wascores = TRUE, expand = TRUE, trace = 1, plot = TRUE)

adonis(dds_dist_N ~ location * treatment * harvest, data = mtx_design, strata = mtx_design$harvest, permutations = 10000)

```
Call:
adonis(formula = dds_dist_N ~ location * treatment * harvest, data = mtx_design,      permutations = 10000, strata = mtx_design$harvest) 

Blocks:  strata 
Permutation: free
Number of permutations: 10000

Terms added sequentially (first to last)

                           Df SumsOfSqs MeanSqs F.Model      R2    Pr(>F)    
location                    1    0.1719 0.17187  1.3497 0.02488   0.07349 .  
treatment                   1    0.5387 0.53865  4.2299 0.07799 9.999e-05 ***
harvest                     3    0.7126 0.23755  1.8654 0.10318   0.00040 ***
location:treatment          1    0.1560 0.15605  1.2254 0.02259   0.15028    
location:harvest            3    0.4812 0.16040  1.2596 0.06967   0.06669 .  
treatment:harvest           3    0.5164 0.17214  1.3517 0.07477   0.02410 *  
location:treatment:harvest  3    0.3825 0.12750  1.0013 0.05538   0.48155    
Residuals                  31    3.9476 0.12734         0.57154              
Total                      46    6.9070                 1.00000              
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1




Use DESEQ2 to compare abundances of viruses in each sample to BULK (no litter)

```{r}
#dds = readRDS(file = paste(working_directory,"/dds_htl_euk_virus.rds", sep = ""))

res = results(dds)
resultsNames(dds)

res = results(dds, contrast = c("group_htl", "H1NoLitterBulk", "H1LitterRhizo"))
comparison = "H1NoLitterBulk_vs_H1LitterRhizo"
res_sig = which(res$padj < 0.05)
length(res_sig)
baseMean_df = data.frame(res$baseMean)
betas_df = data.frame(res$log2FoldChange)
stat_df = data.frame(res$stat)
padj_df = data.frame(res$padj)
setnames(baseMean_df, "res.baseMean", paste0("baseMean_", comparison))
setnames(betas_df, "res.log2FoldChange", paste0("betas_", comparison))
setnames(stat_df, "res.stat",paste0("stat_", comparison))
setnames(padj_df, "res.padj", paste0("padj_", comparison))

res = results(dds, contrast = c("group_htl", "H1NoLitterBulk", "H1LitterBulk"))
comparison = "H1NoLitterBulk_vs_H1LitterBulk"
res_sig = which(res$padj < 0.05)
length(res_sig)
baseMean_df = data.frame(baseMean_df, res$baseMean)
betas_df = data.frame(betas_df, res$log2FoldChange)
stat_df = data.frame(stat_df, res$stat)
padj_df = data.frame(padj_df, res$padj)
setnames(baseMean_df, "res.baseMean", paste0("baseMean_", comparison))
setnames(betas_df, "res.log2FoldChange", paste0("betas_", comparison))
setnames(stat_df, "res.stat",paste0("stat_", comparison))
setnames(padj_df, "res.padj", paste0("padj_", comparison))

res = results(dds, contrast = c("group_htl", "H1NoLitterBulk", "H1NoLitterRhizo"))
comparison = "H1NoLitterBulk_vs_H1NoLitterRhizo"
res_sig = which(res$padj < 0.05)
length(res_sig)
baseMean_df = data.frame(baseMean_df, res$baseMean)
betas_df = data.frame(betas_df, res$log2FoldChange)
stat_df = data.frame(stat_df, res$stat)
padj_df = data.frame(padj_df, res$padj)
setnames(baseMean_df, "res.baseMean", paste0("baseMean_", comparison))
setnames(betas_df, "res.log2FoldChange", paste0("betas_", comparison))
setnames(stat_df, "res.stat",paste0("stat_", comparison))
setnames(padj_df, "res.padj", paste0("padj_", comparison))

res = results(dds, contrast = c("group_htl", "H2NoLitterBulk", "H2LitterRhizo"))
comparison = "H2NoLitterBulk_vs_H2LitterRhizo"
res_sig = which(res$padj < 0.05)
length(res_sig)
baseMean_df = data.frame(baseMean_df, res$baseMean)
betas_df = data.frame(betas_df, res$log2FoldChange)
stat_df = data.frame(stat_df, res$stat)
padj_df = data.frame(padj_df, res$padj)
setnames(baseMean_df, "res.baseMean", paste0("baseMean_", comparison))
setnames(betas_df, "res.log2FoldChange", paste0("betas_", comparison))
setnames(stat_df, "res.stat",paste0("stat_", comparison))
setnames(padj_df, "res.padj", paste0("padj_", comparison))

res = results(dds, contrast = c("group_htl", "H2NoLitterBulk", "H2LitterBulk"))
comparison = "H2NoLitterBulk_vs_H2LitterBulk"
res_sig = which(res$padj < 0.05)
length(res_sig)
baseMean_df = data.frame(baseMean_df, res$baseMean)
betas_df = data.frame(betas_df, res$log2FoldChange)
stat_df = data.frame(stat_df, res$stat)
padj_df = data.frame(padj_df, res$padj)
setnames(baseMean_df, "res.baseMean", paste0("baseMean_", comparison))
setnames(betas_df, "res.log2FoldChange", paste0("betas_", comparison))
setnames(stat_df, "res.stat",paste0("stat_", comparison))
setnames(padj_df, "res.padj", paste0("padj_", comparison))

res = results(dds, contrast = c("group_htl", "H2NoLitterBulk", "H2NoLitterRhizo"))
comparison = "H2NoLitterBulk_vs_H2NoLitterRhizo"
res_sig = which(res$padj < 0.05)
length(res_sig)
baseMean_df = data.frame(baseMean_df, res$baseMean)
betas_df = data.frame(betas_df, res$log2FoldChange)
stat_df = data.frame(stat_df, res$stat)
padj_df = data.frame(padj_df, res$padj)
setnames(baseMean_df, "res.baseMean", paste0("baseMean_", comparison))
setnames(betas_df, "res.log2FoldChange", paste0("betas_", comparison))
setnames(stat_df, "res.stat",paste0("stat_", comparison))
setnames(padj_df, "res.padj", paste0("padj_", comparison))

res = results(dds, contrast = c("group_htl", "H3NoLitterBulk", "H3LitterRhizo"))
comparison = "H3NoLitterBulk_vs_H3LitterRhizo"
res_sig = which(res$padj < 0.05)
length(res_sig)
baseMean_df = data.frame(baseMean_df, res$baseMean)
betas_df = data.frame(betas_df, res$log2FoldChange)
stat_df = data.frame(stat_df, res$stat)
padj_df = data.frame(padj_df, res$padj)
setnames(baseMean_df, "res.baseMean", paste0("baseMean_", comparison))
setnames(betas_df, "res.log2FoldChange", paste0("betas_", comparison))
setnames(stat_df, "res.stat",paste0("stat_", comparison))
setnames(padj_df, "res.padj", paste0("padj_", comparison))

res = results(dds, contrast = c("group_htl", "H3NoLitterBulk", "H3LitterBulk"))
comparison = "H3NoLitterBulk_vs_H3LitterBulk"
res_sig = which(res$padj < 0.05)
length(res_sig)
baseMean_df = data.frame(baseMean_df, res$baseMean)
betas_df = data.frame(betas_df, res$log2FoldChange)
stat_df = data.frame(stat_df, res$stat)
padj_df = data.frame(padj_df, res$padj)
setnames(baseMean_df, "res.baseMean", paste0("baseMean_", comparison))
setnames(betas_df, "res.log2FoldChange", paste0("betas_", comparison))
setnames(stat_df, "res.stat",paste0("stat_", comparison))
setnames(padj_df, "res.padj", paste0("padj_", comparison))

res = results(dds, contrast = c("group_htl", "H3NoLitterBulk", "H3NoLitterRhizo"))
comparison = "H3NoLitterBulk_vs_H3NoLitterRhizo"
res_sig = which(res$padj < 0.05)
length(res_sig)
baseMean_df = data.frame(baseMean_df, res$baseMean)
betas_df = data.frame(betas_df, res$log2FoldChange)
stat_df = data.frame(stat_df, res$stat)
padj_df = data.frame(padj_df, res$padj)
setnames(baseMean_df, "res.baseMean", paste0("baseMean_", comparison))
setnames(betas_df, "res.log2FoldChange", paste0("betas_", comparison))
setnames(stat_df, "res.stat",paste0("stat_", comparison))
setnames(padj_df, "res.padj", paste0("padj_", comparison))

res = results(dds, contrast = c("group_htl", "H4NoLitterBulk", "H4LitterRhizo"))
comparison = "H4NoLitterBulk_vs_H4LitterRhizo"
res_sig = which(res$padj < 0.05)
length(res_sig)
baseMean_df = data.frame(baseMean_df, res$baseMean)
betas_df = data.frame(betas_df, res$log2FoldChange)
stat_df = data.frame(stat_df, res$stat)
padj_df = data.frame(padj_df, res$padj)
setnames(baseMean_df, "res.baseMean", paste0("baseMean_", comparison))
setnames(betas_df, "res.log2FoldChange", paste0("betas_", comparison))
setnames(stat_df, "res.stat",paste0("stat_", comparison))
setnames(padj_df, "res.padj", paste0("padj_", comparison))

res = results(dds, contrast = c("group_htl", "H4NoLitterBulk", "H4LitterBulk"))
comparison = "H4NoLitterBulk_vs_H4LitterBulk"
res_sig = which(res$padj < 0.05)
length(res_sig)
baseMean_df = data.frame(baseMean_df, res$baseMean)
betas_df = data.frame(betas_df, res$log2FoldChange)
stat_df = data.frame(stat_df, res$stat)
padj_df = data.frame(padj_df, res$padj)
setnames(baseMean_df, "res.baseMean", paste0("baseMean_", comparison))
setnames(betas_df, "res.log2FoldChange", paste0("betas_", comparison))
setnames(stat_df, "res.stat",paste0("stat_", comparison))
setnames(padj_df, "res.padj", paste0("padj_", comparison))

res = results(dds, contrast = c("group_htl", "H4NoLitterBulk", "H4NoLitterRhizo"))
comparison = "H4NoLitterBulk_vs_H4NoLitterRhizo"
res_sig = which(res$padj < 0.05)
length(res_sig)
baseMean_df = data.frame(baseMean_df, res$baseMean)
betas_df = data.frame(betas_df, res$log2FoldChange)
stat_df = data.frame(stat_df, res$stat)
padj_df = data.frame(padj_df, res$padj)
setnames(baseMean_df, "res.baseMean", paste0("baseMean_", comparison))
setnames(betas_df, "res.log2FoldChange", paste0("betas_", comparison))
setnames(stat_df, "res.stat",paste0("stat_", comparison))
setnames(padj_df, "res.padj", paste0("padj_", comparison))

res_all_data_bulkcomp = data.frame(betas_df, stat_df, padj_df)
rownames(res_all_data_bulkcomp) = rownames(res)
saveRDS(res_all_data_bulkcomp, file = paste(working_directory,"/res_euk_virus_bulkcomp.rds", sep = ""))

# Only keep p values where the comparison by Wald test was significantly different
padj_matrix = as.matrix(res_all_data_bulkcomp[,25:36]) # This is essentially padj_df, but only saved res_all_data, so working from that
padj_matrix[is.na(padj_matrix)] = 1 # Change NA to 1 so will get filtered out in next step
padj_matrix[which(padj_matrix > 0.05)] = NA # Convert all p values > 0.05 to NA
res_all_data_bulkcomp_sig = data.frame(res_all_data_bulkcomp[,1:24], as.data.frame(padj_matrix)) # Remake res_all_data with only significant p values, all non-significant p values are NA

saveRDS(res_all_data_bulkcomp_sig, file = paste(working_directory,"/res_euk_virus_bulkcomp_sig.rds", sep = ""))
write.table(res_all_data_bulkcomp_sig, file = paste(working_directory,"/results_htl_sig_bulkcomp_euk_virus_020919.txt", sep = ""), append = FALSE)

```


###### Ordination of RNA Euk viruses

```{r}
###### Add path to the experimental design file
designfile = "mtx_design_ecofun_genomes4.txt" 
mtx_design = read.delim(designfile, header = TRUE, row.names = 1)
mtx_design <- mtx_design[!(rownames(mtx_design) %in% c("H3_Rhizo_39")),] # Remove these because these columns are mostly 0s, was messing up the geometric mean 
# Keeping "H2_BulkLitter_50" for now
design_rownames = rownames(mtx_design)

emirge_points = as.data.frame(dds_metaMDS_N$points)
#emirge_metaMDS = emirge_pc
#emirge_points = emirge_pc
#### MAKE SURE SAMPLES ARE IN THE SAME ORDER AS THE POINTS!!!

treatment_column = as.data.frame(mtx_design$loc_treat)
rownames(treatment_column) = rownames(mtx_design)
emirge_points_df = merge(emirge_points, treatment_column, by = "row.names")

emirge_points_B = emirge_points_df[emirge_points_df[,4]=="Bulk_NoLitter",]
emirge_points_BL = emirge_points_df[emirge_points_df[,4]=="Bulk_Litter",]
emirge_points_R = emirge_points_df[emirge_points_df[,4]=="Rhizo_NoLitter",]
emirge_points_RL = emirge_points_df[emirge_points_df[,4]=="Rhizo_Litter",]

emirge_points_B_m = as.matrix(emirge_points_B[,2:3])
emirge_points_BL_m = as.matrix(emirge_points_BL[,2:3])
emirge_points_R_m = as.matrix(emirge_points_R[,2:3])
emirge_points_RL_m = as.matrix(emirge_points_RL[,2:3])

# Make a PCA!

###### Update file name for PDF
pdf(file = paste(working_directory,"/PCA_euk_virus-Narna_nohull_coolcolor.pdf", sep = ""), height=5, width=5,useDingbats=FALSE)
plot(dds_metaMDS_N, display = c("sites"), choices = c(1, 2), type = "p", xlab = " ", ylab = " ")

###### Choose point shape and color
points(emirge_points_BL_m, pch = 21, cex = 1, col="brown", bg = "brown")
points(emirge_points_RL_m, cex = 1, pch = 22, col="tan3", bg = "tan3")
points(emirge_points_B_m, pch = 22, cex = 1, col="blue", bg = "blue")
points(emirge_points_R_m, cex = 1, pch = 21, col="cornflowerblue", bg = "cornflowerblue")

group_factor =  as.factor(emirge_points_df[,4])

#Ellipses around Root vs. Bulk
#Calculate centroids
#centroid_B = summary(ordiellipse(dds_metaMDS, groups = group_factor, show.groups = "Bulk_NoLitter", col = "black", lty = 1, lwd = 1, kind = "se", conf = 0.95))
#centroid_R = summary(ordiellipse(dds_metaMDS, groups = group_factor, show.groups = "Bulk_Litter", col = "black", lty = 1, lwd = 1, kind = "se", conf = 0.95))

#lty 5
#centroid_BL = summary(ordiellipse(dds_metaMDS, groups = group_factor, show.groups = "Rhizo_NoLitter", col = "black", lty = 5, lwd = 1, kind = "se", conf = 0.95))
#centroid_RL = summary(ordiellipse(dds_metaMDS, groups = group_factor, show.groups = "Rhizo_Litter", col = "black", lty = 5, lwd = 1, kind = "se", conf =0.95))

###### Create legend
legend("topleft",legend=c("Rhizo Litter", "Bulk Litter", "Rhizo", "Bulk"),pch= c(22, 21, 21, 22),col= c("tan3", "brown", "cornflowerblue", "blue"), pt.bg = c("tan3", "brown", "cornflowerblue", "blue"), bty = "n")

dev.off()
```


#Treatment

```{r}

harvest_column = as.data.frame(mtx_design$harvest)
rownames(harvest_column) = rownames(mtx_design)
emirge_points_df = merge(emirge_points, harvest_column, by = "row.names")

emirge_points_1 = emirge_points_df[emirge_points_df[,4]=="H1",]
emirge_points_2 = emirge_points_df[emirge_points_df[,4]=="H2",]
emirge_points_3 = emirge_points_df[emirge_points_df[,4]=="H3",]
emirge_points_4 = emirge_points_df[emirge_points_df[,4]=="H4",]

emirge_points_1_m = as.matrix(emirge_points_1[,2:3])
emirge_points_2_m = as.matrix(emirge_points_2[,2:3])
emirge_points_3_m = as.matrix(emirge_points_3[,2:3])
emirge_points_4_m = as.matrix(emirge_points_4[,2:3])

# Make a PCA!

###### Update file name for PDF.  You will need to run all code from "pdf" to "dev.off" to generate PDF
pdf(file = paste(working_directory,"PCA_euk_virus-Narna_nohull_time.pdf", sep = ""), height=5, width=5,useDingbats=FALSE)
plot(dds_metaMDS_N, display = c("sites"), choices = c(1, 2), type = "p", xlab = " ", ylab = " ")

###### change PCH values here
points(emirge_points_1_m, pch = 21, cex = 1, col="yellow", bg = "yellow")
points(emirge_points_2_m, cex = 1, pch = 21, col="brown", bg = "brown")
points(emirge_points_3_m, pch = 21, cex = 1, col="blue", bg = "blue")
points(emirge_points_4_m, cex = 1, pch = 21, col="red", bg = "red")

group_factor =  as.factor(emirge_points_df[,3])

#Ellipses around Root vs. Bulk
#Calculate centroids
#centroid_B = summary(ordiellipse(dds_metaMDS, groups = group_factor, show.groups = "Bulk_NoLitter", col = "black", lty = 1, lwd = 1, kind = "se", conf = 0.95))
#centroid_R = summary(ordiellipse(dds_metaMDS, groups = group_factor, show.groups = "Bulk_Litter", col = "black", lty = 1, lwd = 1, kind = "se", conf = 0.95))

#lty 5
#centroid_BL = summary(ordiellipse(dds_metaMDS, groups = group_factor, show.groups = "Rhizo_NoLitter", col = "black", lty = 5, lwd = 1, kind = "se", conf = 0.95))
#centroid_RL = summary(ordiellipse(dds_metaMDS, groups = group_factor, show.groups = "Rhizo_Litter", col = "black", lty = 5, lwd = 1, kind = "se", conf =0.95))

###### Update pch values for the legend
legend("bottomleft",legend=c("1", "2", "3", "4"),pch= c(21, 22, 21, 22),col= c("yellow", "brown", "blue", "red"), bg = c("yellow", "brown", "blue", "red"), bty = "n")

dev.off()
```